---
title: "R Notebook"
output: html_notebook
---

First, let's get the residents of the municipality The Hague in the year 2006. 

```{r}
year <- 2016
municipality <- "0518"

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
source("cbs_get_residents.R")

res <- get_residents(year, municipality)
```

Now, let's get their household income and personal income. We designed the functions in such a way that
dirs and filters can be overwritten, so make sure not to source two functions right after each other!

```{r}
source("cbs_get_household_income.R")
household_income <- get_household_income(year)

source("cbs_get_personal_income.R")
personal_income <- get_personal_income(year)
```

Let's combine these data tables
```{r}
data <- merge(res, personal_income, by = c("RINPERSOONS", "RINPERSOON"), all.x = TRUE)

#by passing the argument all.x = TRUE, we will keep people who are not associated with a household.
data <- merge(data, household_income, by = c("RINPERSOONSHKW", "RINPERSOONHKW"), all.x = TRUE)
```

Optionally, we can save the created dataset as a parquet file and read that in for future reference. This wil save you a lot of time, since you don't have to wait for each of the previous function to run.
```{r}
filename <- "my_amazing_dataset.parquet"
arrow::write_parquet(data, filename)
# rm(res)
# rm(household_income)
# rm(personal_income)

```
Uncomment the next code chunk if you want to read in the dataset

```{r}
# data <- setDT(arrow::read_parquet(filename))
```

Next, let's just get the primary cause of death for our residents
```{r}
source("cbs_get_prim_cause_of_death.R")
death <- get_prim_cause_of_death(year)
#we don't really care about the year column
death <- death[, 1:3]
```

Let's combine these data tables. Keep in mind that most people will be alive, so keep the all.x = TRUE in tact!
```{r}
data <- merge(data, death, by = c("RINPERSOONS", "RINPERSOON"), all.x = TRUE)
```

Things should be rather straight forward from now! Next up: medication usage, but now I want to have the data for 2014, 2015 and 2016
```{r}
source("cbs_get_med.R")
med <- get_med(c(2014, 2015, 2016))
# let's rename the year column
setnames(med, "year", "medyear")
```

Let's combine these data tables. Keep in mind that most people will be alive, so keep the all.x = TRUE in tact!
```{r}
data <- merge(data, med, by = c("RINPERSOONS", "RINPERSOON"), all.x = TRUE)
```